---
title: "Vinho Verde Wine Analysis"
author: "Evan Jones"
date: "10/2/2020"
output: 
  html_document:
    code_folding: hide
    df_print: paged
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(require(pacman)==F) install.packages("pacman")

pacman::p_load(tidyverse, DataExplorer, caret)

cat("\014")
rm(list = ls())
options(scipen = 10)

source("external_functions/data_summary.R")
```

# Overview

[Explain background of wine process]

[Where dataset comes from]

[What was accomplished]

[What the model is capable of doing]


# Data Description
This dataset contains physicochemical and qualitative measures of red and white vinho verde wine from Portugal. The overall goal of this dataset is to predict wine quality in a 0 to 10 rating scale. Each quality observation is based on an expert review.

There are 13 total variables in the dataset.

[More backstory on where the data is from. Some EU stuff]


Variable Name | Variable Type | Unit | Description
- | - | - | -
Fixed Acidity | Continuous | $g(tartaric\;acid)/dm^3$ | The concentration of tartaric acid in the wine bottle
Volatile Acidity | Continuous | $g(acetic\;acid)/dm^3$ | The concentration of acetic acid in the wine bottle
Citric Acid | Continuous | $g/dm^3$ | The concentration of citric acid in the wine bottle
Residual Sugar | Continuous | $g/dm^3$ | The concentration of sugar in the wine bottle
Chlorides | Continuous | $g/dm^3$ | The concentration of chlorides in the wine bottle
Free Sulfur Dioxide | Discrete | $mg/dm^3$ | The concentration of free sulfur dioxide in the wine bottle
Total Sulfur Dioxide | Discrete | $mg/dm^3$ | The overall concentration of sulfur dioxide in the wine bottle
Density | Continuous | $g/cm^3$ | The liquid density in the wine bottle
pH | Continuous | N/A | A measure of acidity and basicity in an aqueous solution. Values can range between 0 and 14. Values less than 7 indicate an acidic solution, whereas values above 7 indicate a basic solution. Exactly 7 indicates a neutral solution
Sulphates | Continuous | $g(potassium\;sulphate)/dm^3$ | The concentration of sulphates in the wine bottle
Alcohol | Continuous | % volume | % alcohol content of the wine bottle.
Quality | Discrete | N/A | 0-10 score assigned to the bottle
Color | Factor (2 levels) | N/A | Whether the wine is white or red

# Analysis

## Preprocessing

[Something about how code can be expanded]

The following steps were applied in order to prepare the data

1. The most recent copy of the dataset was read in from the UCI archives
```{r, data_read}
# Read in datasets from UCI website

df_white <- read.csv2("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv")
df_red <- read.csv2("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv")

# Local backup for dataset as of 8/23/20

# df_white <- read.csv2("./datasets/winequality-white.csv")
# df_red <- read.csv2("./datasets/winequality-red.csv")

# Union datasets
df_white$color <- "White"
df_red$color <- "Red"

df <- union_all(df_white, df_red)

rm(df_red, df_white)
```


2. Variables were renamed to fit proper naming conventions for snake case
```{r, variable_rename}
# Rename columns to snake case
df <- df %>% 
  rename(fixed_acidity = fixed.acidity,
         volatile_acidity = volatile.acidity,
         citric_acid = citric.acid,
         residual_sugar = residual.sugar,
         free_sulfur_dioxide = free.sulfur.dioxide,
         total_sulfur_dioxide = total.sulfur.dioxide
         )

```


3. Variable types were converted
```{r, type_cast}
# Identify column data types
df.numeric = c("fixed_acidity","volatile_acidity", "citric_acid", "residual_sugar", "chlorides", "density", "pH", "sulphates", "alcohol", "quality")
df.int = c("free_sulfur_dioxide", "total_sulfur_dioxide")
df.factor = c("color")

df[df.numeric] <- lapply(df[df.numeric], as.numeric)
df[df.int] <- lapply(df[df.int], as.integer)
df[df.factor] <- lapply(df[df.factor], as.factor)

rm(df.numeric, df.int, df.factor)

```

4. The wine color was releveled so that white wine would act as the base level
```{r}
# relevel color base as white (red wines are generally more premium)
df$color <- df$color %>% relevel(ref = "White")
```


## Exploratory Analysis

**Amount of Missing values**

There are no missing values


```{r, exploratory_plots}
# Exploratory Data Analysis Plots
# plot_intro(df)
plot_bar(df)
plot_histogram(df)

```

**Multicollinearity Check**
```{r, correlation}
plot_correlation(df, type = "c")
```

Free sulfur dioxide and total sulfur dioxide are highly correlated so only one of the two variables was used in the final model


From this we notice that free and total sulfur dioxide are highly correlated. Thus only one of the two was used in the analysis.

**Outlier Investigation**

```{r, data_summary, include = F}
data.summary(df)
```

```{r, outlier_plots, figures-side, fig.show="hold", out.width="50%"}
# Investigate unusual columns
plot(df$quality, df$density,
     ylab = "Density", 
     xlab = "Quality", 
     main = "Quality vs. Density",
     col=ifelse(df$density > 1.03, "red", "black"))

plot(df$quality, df$residual_sugar,
     ylab = "Residual Sugar", 
     xlab = "Quality", 
     main = "Quality vs. Residual Sugar",
     col=ifelse(df$residual_sugar > 60, "red", "black"))

```
```{r, outlier_removal, include=F}
# Remove outliers from residual sugar and density

df_outlier <- which(df$residual_sugar > 60)

df_outlier %>% 
  union(which(df$density > 1.03)) -> df_outlier

df = df[-df_outlier,]

```

## Final Explanatory Model
```{r, training_split}
# Set random seed and split data into training and validation sets
set.seed(3)
trainIndex = sample(1:nrow(df), size = round(0.6*nrow(df)), replace=FALSE)
df.train <- df[trainIndex, ]
df.valid <- df[-trainIndex, ]
rm(trainIndex)
```

For the final model, training and validation datasets were constructed. 60% of the data was used for training, and 40% of the data was used for validation.

```{r, model}
# Linear model
lm(sqrt(quality) ~
  volatile_acidity +
  free_sulfur_dioxide +
  alcohol +
  color,
  data = df.train) -> df.reg

summary(df.reg)
```

[Some commentary about overall model size]

The variables used in the final model are:

* Volatile acidity
* Free sulfur dioxide
* Alcohol
* Color

[Model formula]

The $R^{2}_{adj}$ for the model was [some value]

```{r, rmse}
# Calculate predicted values for validation set and transform
df.valid$quality_pred <- predict(df.reg, newdata=df.valid)
df.valid <- df.valid %>% 
  mutate(quality_pred_adj = I(quality_pred^2))

# Compare RMSE calculation to caret package
rmse <- caret::RMSE(df.valid$quality_pred_adj, df.valid$quality)
```

The model has an overall RMSE of `r round(rmse, digits=2)`

### Model Assumptions

**Mean of Error Term**

```{r, include = F}
mean(df.reg$residuals)
```

The mean of the probability distribution of the error is $\approx 0$, Therefore this assumption is passed.

**Variance of Error Term**

```{r}
plot(df.reg$fitted.values, df.reg$residuals, 
     ylab = "Residuals", 
     xlab = "Fitted Values", 
     main = "Residuals vs. Fitted Values")
abline(0,0, col = "red")
```

The variance of the probability distribution of the error looks [insert], so this assumptions is passed.

**Normal Distribution of Error Term**

```{r, qqplot}
qqnorm(df.reg$residuals)
qqline(df.reg$residuals, col = "red")
```

The QQ Plot deviation in the lower left and upper right corners indicates long tails on both the lower and upper ends of the wine quality.

Between this plot and the above residual plot, we can assume that the error is normally distributed.

**Independent Observations**

Additionally, we can assume that each of these observations are independent of each other as each one represents a separate bottle of wine.

### Predictive Power

Since there are only 4 observations with a quality of 9, and no observations with a quality of 10, high predicted qualities should be approached with caution.

**Example Prediction 1 - White Wine**
```{r, predict1}

```

The first example prediction was made on a bottle of wine with the following characteristics:

[characteristics]

**Example Prediction2 - Red Wine**
```{r, predict2}

```

The second example prediction was made on a bottle of wine with the following characteristics:

[characteristics]

# Conclusion

[Address business value proposition]

[Address how solution meets requirements of customer segment]
