---
title: "analytics_workflow"
author: "Evan Jones"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(require(pacman)==F) install.packages("pacman")

pacman::p_load(tidyverse, DataExplorer, caret)

cat("\014")
rm(list = ls())
options(scipen = 10)

source("external_functions/data_summary.R")
```

# Overview

# Data Description
This dataset contains physicochemical and qualitative measures of red and white vinho verde wine from Portugal. The overall goal of this dataset is to predict wine quality in a 0 to 10 rating scale. Each quality observation is based on an expert review.

There are 13 total variables in the dataset.


Variable Name | Variable Type | Unit | Description
- | - | - | -
Fixed Acidity | Continuous | $g(tartaric\;acid)/dm^3$ | The concentration of tartaric acid in the wine bottle
Volatile Acidity | Continuous | $g(acetic\;acid)/dm^3$ | The concentration of acetic acid in the wine bottle
Citric Acid | Continuous | $g/dm^3$ | The concentration of citric acid in the wine bottle
Residual Sugar | Continuous | $g/dm^3$ | The concentration of sugar in the wine bottle
Chlorides | Continuous | $g/dm^3$ | The concentration of chlorides in the wine bottle
Free Sulfur Dioxide | Discrete | $mg/dm^3$ | The concentration of free sulfur dioxide in the wine bottle
Total Sulfur Dioxide | Discrete | $mg/dm^3$ | The overall concentration of sulfur dioxide in the wine bottle
Density | Continuous | $g/cm^3$ | The liquid density in the wine bottle
pH | Continuous | N/A | A measure of acidity and basicity in an aqueous solution. Values can range between 0 and 14. Values less than 7 indicate an acidic solution, whereas values above 7 indicate a basic solution. Exactly 7 indicates a neutral solution
Sulphates | Continuous | $g(potassium\;sulphate)/dm^3$ | The concentration of sulphates in the wine bottle
Alcohol | Continuous | % volume | % alcohol content of the wine bottle.
Quality | Discrete | N/A | 0-10 score assigned to the bottle
Color | Factor (2 levels) | N/A | Whether the wine is white or red

# Preprocessing
```{r, include = F}
# Read in datasets from UCI website

df_white <- read.csv2("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv")
df_red <- read.csv2("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv")

# Local backup fo dataset as of 8/23/20

# df_white <- read.csv2("./datasets/winequality-white.csv")
# df_red <- read.csv2("./datasets/winequality-red.csv")

# Union datasets
df_white$color <- "White"
df_red$color <- "Red"

df <- union_all(df_white, df_red)

rm(df_red, df_white)
```

```{r, include=F}
# Rename columns to snake case
df <- df %>% 
  rename(fixed_acidity = fixed.acidity,
         volatile_acidity = volatile.acidity,
         citric_acid = citric.acid,
         residual_sugar = residual.sugar,
         free_sulfur_dioxide = free.sulfur.dioxide,
         total_sulfur_dioxide = total.sulfur.dioxide
         )

```

```{r, include=F}
# Identify column data types
df.numeric = c("fixed_acidity","volatile_acidity", "citric_acid", "residual_sugar", "chlorides", "density", "pH", "sulphates", "alcohol", "quality")
df.int = c("free_sulfur_dioxide", "total_sulfur_dioxide")
df.factor = c("color")

df[df.numeric] <- lapply(df[df.numeric], as.numeric)
df[df.int] <- lapply(df[df.int], as.integer)
df[df.factor] <- lapply(df[df.factor], as.factor)

rm(df.numeric, df.int, df.factor)

# relevel color base as white (red wines are generally more premium)

df$color <- df$color %>% relevel(ref = "White")

str(df)

```

# Analysis

Free sulfur dioxide and total sulfur dioxide are highly correlated so only one of the two variables was used in the final model


```{r, include=F}
# Data Explorer Generation
# DataExplorer::create_report(df)

# report_config <- configure_report(
#   add_introduce = TRUE,
#   add_plot_intro = TRUE,
#   # add_plot_str = TRUE,
#   add_plot_missing = TRUE,
#   add_plot_histogram = TRUE,
#   add_plot_density = FALSE,
#   add_plot_qq = TRUE,
#   add_plot_bar = TRUE,
#   add_plot_correlation = TRUE,
#   # add_plot_prcomp = TRUE,
#   add_plot_boxplot = TRUE,
#   add_plot_scatterplot = TRUE,
#   introduce_args = list(),
#   plot_intro_args = list(),
#   plot_str_args = list(type = "diagonal", fontSize = 35, width = 1000, margin =
#     list(left = 350, right = 250)),
#   plot_missing_args = list(),
#   plot_histogram_args = list(),
#   plot_density_args = list(),
#   plot_qq_args = list(sampled_rows = 1000L),
#   plot_bar_args = list(),
#   plot_correlation_args = list(cor_args = list(use = "pairwise.complete.obs")),
#   plot_prcomp_args = list(),
#   plot_boxplot_args = list(),
#   plot_scatterplot_args = list(sampled_rows = 1000L),
#   global_ggtheme = quote(theme_gray()),
#   global_theme_config = list()
# )
# 
# create_report(
#   df,
#   output_dir = "output/",
#   y = "quality",
#   report_title = "Wine Summary",
#   config = report_config
# )

```

```{r, echo=F}
# Exploratory Data Analysis Plots
plot_intro(df)
plot_bar(df)
plot_histogram(df)
plot_correlation(df, type = "c")
```

From this we notice that free and total sulfur dioxide are highly correlated. Thus only one of the two was used in the analysis.

```{r, include = F}
data.summary(df)
```

```{r, include=F}
# Investigate unusual columns
plot(df$quality, df$density)
plot(df$quality, df$residual_sugar)
plot(df$quality, df$sulphates)
plot(df$quality, df$total_sulfur_dixoide)
plot(df$quality, df$volatile_acidity)
plot(df$quality, df$pH)

```

```{r}
# Investigate outliers

# outliers include
# density, residual sugar, sulphates, total sulfur dioxide, volatile acidity

# outlier at 1.04ish density and over 60 residual


# df_outlier = df[df$residual_sugar > 60,]
# df_outlier

df_outlier = which(df$residual_sugar > 60)
df_outlier2 = which(df$density > 1.03)

```

```{r}
# Remove outliers and replot

df = df[-df_outlier,]

plot(df$quality, df$density)
```

```{r, include=F}
plot(df$quality, df$density)
plot(df$quality, df$residual_sugar)
plot(df$quality, df$sulphates)
plot(df$quality, df$total_sulfur_dixoide)
plot(df$quality, df$volatile_acidity)
```

## Model Building
```{r, include=F}
set.seed(3)
trainIndex = sample(1:nrow(df), size = round(0.6*nrow(df)), replace=FALSE)
df.train <- df[trainIndex, ]
df.valid <- df[-trainIndex, ]
rm(trainIndex)
```

```{r, include=F}
lm(sqrt(quality) ~
  volatile_acidity +
  free_sulfur_dioxide +
  alcohol +
  color,
  data = df.train) -> df.reg

summary(df.reg)
```

```{r, include=F}
plot(df.reg)
```


# Conclusion
