---
title: "analytics_workflow"
author: "Evan Jones"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(require(pacman)==F) install.packages("pacman")

pacman::p_load(tidyverse, DataExplorer, caret)

cat("\014")
rm(list = ls())
options(scipen = 10)
```

# Overview

# Data Description
This dataset contains physicochemical and qualitative measures of red and white vinho verde wine from Portugal. The overall goal of this dataset is to predict wine quality in a 0 to 10 rating scale. Each quality observation is based on an expert review.

There are 13 total variables in the dataset.


Variable Name | Variable Type | Unit | Description
- | - | - | -
Fixed Acidity | Continuous | $g(tartaric\;acid)/dm^3$ | The concentration of tartaric acid in the wine bottle
Volatile Acidity | Continuous | $g(acetic\;acid)/dm^3$ | The concentration of acetic acid in the wine bottle
Citric Acid | Continuous | $g/dm^3$ | The concentration of citric acid in the wine bottle
Residual Sugar | Continuous | $g/dm^3$ | The concentration of sugar in the wine bottle
Chlorides | Continuous | $g/dm^3$ | The concentration of chlorides in the wine bottle
Free Sulfur Dioxide | Discrete | $mg/dm^3$ | The concentration of free sulfur dioxide in the wine bottle
Total Sulfur Dioxide | Discrete | $mg/dm^3$ | The overall concentration of sulfur dioxide in the wine bottle
Density | Continuous | $g/cm^3$ | The liquid density in the wine bottle
pH | Continuous | N/A | A measure of acidity and basicity in an aqueous solution. Values can range between 0 and 14. Values less than 7 indicate an acidic solution, whereas values above 7 indicate a basic solution. Exactly 7 indicates a neutral solution
Sulphates | Continuous | $g(potassium\;sulphate)/dm^3$ | The concentration of sulphates in the wine bottle
Alcohol | Continuous | % volume | % alcohol content of the wine bottle.
Quality | Discrete | N/A | 0-10 score assigned to the bottle
Color | Factor (2 levels) | N/A | Whether the wine is white or red

# Preprocessing
```{r, include = F}
# Read in datasets

df_white <- read.csv2("./datasets/winequality-white.csv")
df_red <- read.csv2("./datasets/winequality-red.csv")

# Union datasets
df_white$color <- "White"
df_red$color <- "Red"

df <- union_all(df_white, df_red)

rm(df_red, df_white)

# Rename columns to snake case
df <- df %>% 
  rename(fixed_acidity = fixed.acidity,
         volatile_acidity = volatile.acidity,
         citric_acid = citric.acid,
         residual_sugar = residual.sugar,
         free_sulfur_dioxide = free.sulfur.dioxide,
         total_sulfur_dioxide = total.sulfur.dioxide
         )

```

```{r}
df.numeric = c("fixed_acidity","volatile_acidity", "citric_acid", "residual_sugar", "chlorides", "density", "pH", "sulphates", "alcohol", "quality")
df.int = c("free_sulfur_dioxide", "total_sulfur_dioxide")
df.factor = c("color")

df[df.numeric] <- lapply(df[df.numeric], as.numeric)
df[df.int] <- lapply(df[df.int], as.integer)
df[df.factor] <- lapply(df[df.factor], as.factor)

rm(df.numeric, df.int, df.factor)

df$color <- df$color %>% relevel(ref = "White")

str(df)

```


# Analysis

Free sulfur dioxide and total sulfur dioxide are highly correlated so only one of the two variables was used in the final model

```{r}
# plot_intro(df)
plot_bar(df)
plot_histogram(df)
plot_correlation(df, type = "c")
```
```{r}
hist(df$residual_sugar, breaks = 100)
# DataExplorer::create_report(df)
```


## Model Building
```{r}
set.seed(3)
trainIndex = sample(1:nrow(df), size = round(0.6*nrow(df)), replace=FALSE)
df.train <- df[trainIndex, ]
df.valid <- df[-trainIndex, ]
rm(trainIndex)
```

```{r}
df.reg.full <- lm(log(quality)~., data = df.train)

summary(df.reg.full)

# lm(quality ~
#   fixed_acidity +
#   volatile_acidity +
#   citric_acid +
#   residual_sugar +
#   chlorides +
#   free_sulfur_dioxide +
#   total_sulfur_dioxide +
#   density +
#   pH +
#   sulphates +
#   alcohol +
#   color,
#   data = df.train) -> df.reg.full2
# 
# 
# summary(df.reg.full2)

```

```{r}
# plot(df.reg.full)

plot(df.reg.full$fitted.values, df.reg.full$residuals, 
     ylab = "Residuals", 
     xlab = "Casual", 
     main = "Residuals vs. Fitted Values")
abline(0,0, col = "red")

qqnorm(df.reg.full$residuals)
qqline(df.reg.full$residuals, col = "red")
```



# Conclusion
