---
title: "Wine Quality Dataset Description"
author: "Evan Jones"
date: "8/23/20"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

```{r setup, include = F}
if(require(pacman)==F) install.packages("pacman")

pacman::p_load(tidyverse, DataExplorer, snakecase, gdata)

cat("\014")
rm(list = ls())
options(scipen = 10)
```

```{r, include = F}
# Read in datasets

df_white <- read.csv2("./datasets/winequality-white.csv")
df_red <- read.csv2("./datasets/winequality-red.csv")

gdata::combine(
  df_white,
  df_red,
  names = c("W", "R")
) -> df

rm(df_red, df_white)

colnames(df) %>% 
  snakecase::to_lower_camel_case() -> colnames(df)

colnames(df)[colnames(df) == "source"] <- "color"

```

# Dataset Overview

This dataset contains physicochemical and qualitative measures of red and white vinho verde wine from Portugal. The overall goal of this dataset is to predict wine quality in a 0 to 10 rating scale. Each quality observation is based on an expert review.

## Number of Observations
```{r}
nrow(df)
df %>% count(color)
```
This dataset contains 6,497 observations. Out of this total, 4,898 are observations are for white wine and 1,599 are for red wine.

## Variable Overview
```{r, echo = F}
str(df)
```
There are 13 total variables in the dataset.

Variable Name | Variable Type | Unit | Description
- | - | - | -
Fixed Acidity | Continuous | $g(tartaric\;acid)/dm^3$ | The concentration of tartaric acid in the wine bottle
Volatile Acidity | Continuous | $g(acetic\;acid)/dm^3$ | The concentration of acetic acid in the wine bottle
Citric Acid | Continuous | $g/dm^3$ | The concentration of citric acid in the wine bottle
Residual Sugar | Continuous | $g/dm^3$ | The concentration of sugar in the wine bottle
Chlorides | Continuous | $g/dm^3$ | The concentration of chlorides in the wine bottle
Free Sulfur Dioxide | Discrete | $mg/dm^3$ | The concentration of free sulfur dioxide in the wine bottle
Total Sulfur Dioxide | Discrete | $mg/dm^3$ | The overall concentration of sulfur dioxide in the wine bottle
Density | Continuous | $g/cm^3$ | The liquid density in the wine bottle
pH | Continuous | N/A | A measure of acidity and basicity in an aqueous solution. Values can range between 0 and 14. Values less than 7 indicate an acidic solution, whereas values above 7 indicate a basic solution. Exactly 7 indicates a neutral solution
Sulphates | Continuous | $g(potassium\;sulphate)/dm^3$ | The concentration of sulphates in the wine bottle
Alcohol | Continuous | % volume | % alcohol content of the wine bottle.
Quality | Discrete | N/A | 0-10 score assigned to the bottle
Color | Factor (2 levels) | N/A | Whether the wine is white or red

## Data Completeness
```{r}
plot_intro(df)
```

This dataset is considered complete with no missing values for any observation.

## Planned Analysis

For this analysis, I first plan on conducting heavy background research to have a better idea for the context of this dataset. This would start with looking at the vinho verde wine. I would want to see if it is a protected designation of wine. This is similar to how Italy protects many of their agricultural exports with special designations of quality. There could be some sort of scientific indicator for what constitutes a quality vinho verde wine.

This would then lead me to research each variable to better understand the underlying chemistry behind wine observations. I would want to know how tartaric acid, acetic acid, citric acid, sugar, chlorides, free and total sulfur dioxide content, overall density, pH, sulphate content, and alcohol content contribute to a higher quality wine and in what amounts.

I plan on reading this article: https://onlinelibrary.wiley.com/doi/10.1002/9781118730720.ch3

I would then feel comfortable enough to complete a more thorough exploratory data analysis. I would want to potentially reduce the dimensionality of the dataset if any of the variables are too closely related. Then I would choose an appropriate modeling method to construct a model. I may want to break out the data into red wine and white wine subsets and construct separate models for each.

Once I feel my models accurately reflect the ability to predict vinho verde wine quality, I will provide some example predictions and visualizations of the model prediction power. Finally, I will explain any assumptions or model shortcomings.